Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-07-02 manuscript/manuscript.Rmd
Latitude

Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams.
Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:penguin-body-mass}:
no transformation (residual sum $\sum r_i = 48.0$);
conventional transformation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.102$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
shift- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
and robust shift- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).

@@

out <- .plot_penguin_body_mass(plot_theme = plot_theme)
out$plot
```
Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass variable after transformation.
The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).
The distribution of the original data was not centrally normal ($p = 0.03$).

### Longitudinal coordinates in the Ames housing dataset

Datasets where invariance has better properties in terms of lambda and resulting mean and standard deviation.
modeldata::ischemic_stroke$age
modeldata::ames$Longitudinal

>>>>

Gorman, Williams and Fraser recorded body mass (in grams) of 342 penguins of three different species [@Gorman2014-eo].
The body mass was $(4.2 \pm 0.8) \cdot 10^3$ (mean ± standard deviation) grams, and not centrally normal ($p = 0.03$).
Applying conventional and invariant Yeo-Johnson transformations to patient age yielded the following results, see Figure \ref{fig:penguin-body-mass}:
no transformation (residual sum $\sum r_i = 48.0$);
conventional transformation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.102$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.5$, $\sum r_i = 32.2$, $\mu_{YJ} = 2.101$, $\sigma_{YJ} = 4 \cdot 10^{-3}$);
shift- and scale invariant transformation ($\lambda = 0.5$, $\sum r_i = 26.8$, $\mu_{YJ} = 0.9$, $\sigma_{YJ} = 0.9$);
and robust shift- and scale-invariant transformation ($\lambda = 0.4$, $\sum r_i = 23.1$, $\mu_{YJ} = 0.8$, $\sigma_{YJ} = 0.8$).

@@

out <- .plot_penguin_body_mass(plot_theme = plot_theme)
out$plot
```
Shift- and scale-invariant transformation leads to a lower overall residual sum, indicating a better transformation.
Moreover, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass variable after transformation.
The empirical central normality test did not detect any statistically significant deviations from central normality for any transformation (all $p \geq 0.38$).

### Latitude in the Ames housing dataset

Geospatial datasets usually contain coordinates.
The Ames housing dataset contains data on 2930 properties that were sold between 2006 and 2010 [@De-Cock2011-jf], including their geospatial coordinates.
The latitude was $42.03 \pm 0.02)$ (mean ± standard deviation).
Applying conventional and invariant Yeo-Johnson transformations to latitude yielded the following results, see Figure \ref{fig:ames-latitude}:
no transformation (residual sum $\sum r_i = 328$);
conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{YJ} = 4.8 \cdot 10^{99}$, $\sigma_{YJ} = 0.1 \cdot 10^{99}$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 95.4$, $\sum r_i = 319$, $\mu_{YJ} = $, $\sigma_{YJ} = $);
shift- and scale invariant transformation ($\lambda = 1.5$, $\sum r_i = 326$, $\mu_{YJ} = $, $\sigma_{YJ} = $);
and robust shift- and scale-invariant transformation ($\lambda = 1.4$, $\sum r_i = 311$, $\mu_{YJ} = $, $\sigma_{YJ} = $).

```{r ames-latitude, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Quantile-quantile plots for penguin body mass, measured in grams. ",
  "The left panel shows the quantile-quantile plot for the original data. ",
  "The middle panel shows the quantile-quantile plots for data transformed using ",
  "the original Yeo-Johnson transformation, and Raymaekers and Rousseeuw's robust adaptation \\cite{Raymaekers2021-kq}. ",
  "The right panels shows the quantile-quantile plots for data transformed using ",
  "the non-robust and robust invariant transformations."
)

out <- .plot_ames_latitude(plot_theme = plot_theme)
out$plot
```
